---
title: "Data Cleaning"
author: "Mireia Valle"
date: "19/2/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading (or installing if needed) the packages we are going to use
```{r, warning = FALSE, message = FALSE}
packages <- c("tidyverse", 
              "tidyr")

for (package in packages) { if (!(package %in% installed.packages())) { install.packages(package) } }

library(tidyverse)
#install.packages("rlang")
library(rlang)
#install.packages("tidyr") #we need the newer version of this package
library(tidyr)
```

## Read in data

We are reading data from [Mike Byerly. Alaska commercial salmon catches by management region (1886- 1997). Gulf of Alaska Data Portal. df35b.304.2.](https://knb.ecoinformatics.org/view/df35b.304.2)

```{r}
catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"), stringsAsFactors = F)
```

*the argument stringsAsFactors = F, ask to deal with strings without transforming to factors. 

```{r}
head(catch_original)
```

The pipe %>% ctrl + shift + M

## Clean and reshape

Remove marginal sum and notes using "select()"

```{r}
catch_clean <- catch_original %>% 
  #select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum) 
  select(-All, -notesRegCode)

head(catch_clean)

summary(catch_clean)

#Chinook has been read as character. We need to change that:

catch_clean <- catch_original %>% 
  #select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum) 
  select(-All, -notesRegCode) %>% 
  mutate(Chinook = as.numeric(Chinook))#overwrite the Chinook column

#we get a warning message: NAs introduced by coercion. Something need to be cleaned in that column because it can't be transformed to numeric
```

## QA quality check

Find where as.numeric couldn't convert Chinook to a number

```{r}
i <- which(is.na(catch_clean$Chinook))
i
```

We now know that the error is in line 401. Let's see the content of that line in the original dataset. 

```{r}
catch_original[i,] #it shows the complete info on that row
```

We see that Chinook has I value and we assume that it should be a 1. 

```{r}
catch_clean <- catch_original %>% 
  #select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum) 
  select(-All, -notesRegCode) %>% 
  mutate(Chinook = ifelse(Chinook=="I", 1, Chinook)) %>% #in case you find that Chinook's value is I, then write 1, if you don't find a I, then type the value you find
  mutate(Chinook = as.numeric(Chinook))#overwrite the Chinook column
```

Reshape from wide to tall format using 'pivot_longer'

```{r}
catch_long <- catch_clean %>% 
  pivot_longer(cols = -c(Region, Year), 
               names_to = "species", 
               values_to = "catch")
head(catch_long)
```

Example with pivot_wider

```{r}
catch_wide <- catch_long %>% 
  pivot_wider(names_from = Year, 
              values_from = catch)

head(catch_wide)
```

Continue reshaping, we correct catch value

```{r}
catch_long <- catch_clean %>% 
  pivot_longer(cols = -c(Region, Year), 
               names_to = "species", 
               values_to = "catch") %>%
  rename(catch_thousands = catch) %>% #rename the column catch
  mutate(catch = catch_thousands * 1000) %>% #we create a new column that contains the value of catch on thousands
  select(-catch_thousands) #we remove the catch_thousands column we created       #mutate(catch = catch * 1000) #using mutate we will make the same but faster overwriten the catch column
```

## group_by and summarize

Calculate the mean catch per region
```{r}
mean_region <- catch_long %>% 
  group_by(Region) %>% #it adds the grouping information
  summarise(catch_mean = mean(catch),
            num_obs = n())#it will generate a new column with the mean value
```

Grouping by multiple variables

```{r}
mean_region_species <- catch_long %>% 
  group_by(Region, species) %>% #it adds the grouping information
  summarise(catch_mean = mean(catch),
            num_obs = n())#it will generate a new column with the mean value and the number of species
```

A species table
```{r}
mean_s <- catch_long %>% 
  group_by(species) %>% #it adds the grouping information
  summarise(catch_mean = mean(catch),
            num_obs = n())
```


The highest salmon catch using summarise ()
```{r}
max_catch <- catch_long %>%
  group_by(Year) %>% 
   summarise(max_catch = max(catch),
             total_catch = sum(catch))
```

```{r}
annual_catch <- catch_long %>% 
  group_by(Year) %>% 
  summarise(total_catch = sum(catch)) %>% 
  arrange(-total_catch)
```

What year has the highest Sockeye catch

```{r}
annual_sockeye <- catch_long %>% 
  filter(species == "Sockeye") %>% 
  group_by(Year) %>% 
  summarise(total_catch = sum(catch)) %>% 
  arrange(-total_catch)

head(annual_sockeye)
```

## Join to 

Reading the data on Regions information

```{r}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", method = "libcurl"), stringsAsFactors = F)

head(region_defs)
```

```{r}
catch_joined <- left_join(catch_long, region_defs, by = c("Region" = "code"))

head(catch_joined)

#using %>% 

#catch_joined <- catch_long %>% 
#  left_join(region_defs)

#or we can rename the columns

region_defs <- region_defs %>% 
  rename(Region = code)

#and then join

catch_joined <- left_join(catch_long, region_defs, by = "Region")

head(catch_joined)
```

