---
title: "Day3_data_visualization"
author: "Mireia Valle"
date: "20/2/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading all libraries
```{r, warning = FALSE, message = FALSE}
library(leaflet)
library(dplyr)
library(tidyr)
library(ggplot2)
library(DT)
library(scales) # install.packages("scales")
library(patchwork) # install.packages("patchwork")
```

## Loading data

```{r}
data_url <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e"

# we use tryCatch to look if the data we want to load is in our project project, in case it isn't, then we will get an error message and it will download it from the url we gave above

esc <- tryCatch(
    read.csv("data/escapement.csv", stringsAsFactors = FALSE),
    error=function(cond) {
        message(paste("Escapement file does not seem to exist, so get it from the KNB."))
        esc <- read.csv(url(data_url, method = "libcurl"), stringsAsFactors = FALSE)
        return(esc)
    }
)

head(esc)

```

## Data Processing

We want to calculate the annual escapment by species and region

```{r}
annual_esc <- esc %>% 
  #we are going to break the date column into three different columns
  separate(sampleDate, c("Year", "Month", "Day"), sep = "-") %>% 
  #transform the new columns we created to numeric 
  mutate(Year = as.numeric(Year),
         Month = as.numeric(Month),
         Day = as.numeric(Day)) %>%
  #Group by multiple columns, it creates subsets of your data
  group_by(Species, SASAP.Region, Year) %>% 
  #calculate statistics
  summarise(escapement = sum(DailyCount)) %>% 
  #we don't want all species, we only want some of them
  filter(Species %in% c("Chinook", "Sockeye", "Chum", "Coho", "Pink"))

head(annual_esc)
```

Example using %in% operator
```{r}
v <- c(1,2,3)

2 %in% v
```

## Data Visualization

### First look at the data

Making a bar plot

```{r}
ggplot(annual_esc, aes(x = Species, y = escapement)) +
  geom_col()
```

Changing the color of the geometric's fill to blue

```{r}
ggplot(annual_esc, aes(x = Species, y = escapement)) +
  geom_col(fill = "blue")
```

Distinguish each region

```{r}
ggplot(annual_esc, aes(x = Species, y = escapement, fill = SASAP.Region)) +
  geom_col()
```

### Looking into the Kodiak region

```{r}
annual_esc %>% 
  filter(SASAP.Region == "Kodiak") %>% 
  ggplot(aes(x = Year, y = escapement)) +
  geom_line()
```

What is going on for each species

```{r}
annual_esc %>% 
  filter(SASAP.Region == "Kodiak") %>% 
  ggplot(aes(x = Year, y = escapement, color = Species)) +
  geom_line() +
  #adding points to observations
  geom_point() +
  #adding a label
  ylab("Escapment") +
  #adding a title
  ggtitle("Salmon Annual Escapement for Kodiak Region") +
  #making the background white
  theme_bw() +
  #center the title
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom")

```

Save your theme to an object

```{r}
my_theme <-   theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom",
        #removing the title of the legend
        legend.title = element_blank())
```

Using my_theme

```{r}
annual_esc %>% 
  filter(SASAP.Region == "Kodiak") %>% 
  ggplot(aes(x = Year, y = escapement, color = Species)) +
  geom_line() +
  #adding points to observations
  geom_point() +
  #adding a label
  ylab("Escapment") +
  #adding a title
  ggtitle("Salmon Annual Escapement for Kodiak Region") +
  #using my_theme
  my_theme
```

Playing with Facet, we are going to plot the escapement data for each region as a multiple plot

```{r}
ggplot(annual_esc, aes(x = Year, y = escapement, color = Species)) + 
    geom_line() +
    geom_point() +
  #set the labels to comma
    scale_y_continuous(labels = comma) +
  #calling the facet
    facet_wrap(~SASAP.Region, scales = "free_y", ncol = 2) +
    ylab("Escapement") +
    my_theme
```

Using patchwork for plotting plots together

```{r}
#We create an object with the ggplot arguments
test_ggplot <- ggplot(annual_esc, aes(x = Species, y = escapement)) +
  geom_col()

test_ggplot

test_ggplot2 <- ggplot(annual_esc, aes(x = Species, y = escapement)) +
  geom_col(fill = "blue")

test_ggplot2

#ploting the two plots together with patchwork
test_ggplot + test_ggplot2

#one after the other
test_ggplot / test_ggplot2

```

