---
title: "Day4_SpatialAnalysis"
author: "Mireia Valle"
date: "21/2/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading libraries

```{r}
library(sf)
library(dplyr)
library(ggplot2)
library(scales)
library(ggmap)
library(leaflet)
```

## Read in data

```{r}
ak_regions <- read_sf("data/ak_regions_simp.shp")

#ploting with base R
plot(ak_regions)# something is happening realted to the coordinate reference system

#what is ak_regions?
class(ak_regions)

#it is a dataframe so we can call head()
head(ak_regions)

#to see the coordinate system
st_crs(ak_regions)
```

### Transforming coordinate systems

```{r}
ak_regions_3338 <- ak_regions %>% 
  st_transform(crs = 3338)

plot(ak_regions_3338)
```

## We can use tidyverse functions to the sf object
```{r}
ak_regions_3338 %>%
  filter(region == "Southeast") %>% #select rows
  select(region) # select column
```

## EPSG CODES
epsg.io interesting link

3338- Alaska Albers
4326- WGS84 (GPS)
3847- pseudo mercator (Google maps, Open street maps)

## Spatial joins and summaries

```{r}
pop <- read.csv("data/alaska_population.csv", stringsAsFactors = F)

head(pop)

class(pop)#it is a data frame and we need a sf object

#we are going to convert a data frame to a sf

pop_4326 <- st_as_sf(pop,
                     coords = c("lng", "lat"), # first is x and then y
                     crs = 4326,
                     remove = F) #It keeps the lat long coordinates we originally have 
```

### Making the join

```{r}
#pop_joined <- st_join(pop_4326, ak_regions_3338, join = st_within)
```

we get an error because the projections are different

We transform our pop 
```{r}
pop_3338 <- pop_4326 %>% 
  st_transform(crs = 3338)
```

We join 
```{r}
pop_joined <- st_join(pop_3338, ak_regions_3338, join = st_within)
```

We plot it
```{r}
plot(pop_joined)
```

### we are going to calculate population by region

```{r}
pop_region <- pop_joined %>% 
  group_by(region) %>% 
  summarise(total_pop = sum(population))
```

### droping geometry column. For help: ?sf::tidyverse

```{r}
pop_region <- pop_joined %>% 
  as.data.frame() %>% 
  group_by(region) %>% 
  summarise(total_pop = sum(population))
```

### now we join the geometry from ak_regions to pop_region 

```{r}
pop_region_3338 <- left_join(ak_regions_3338, pop_region, by = "region")

plot(pop_region_3338)
```

### calculation total population on management area (mgmt column)

```{r}
pop_mgmt_3338 <- pop_region_3338 %>% 
  group_by(mgmt_area) %>% 
  summarise(total_pop = sum(total_pop), do_union = FALSE) #do.union= F keeps the lines of the regions

plot(pop_mgmt_3338["total_pop"])
```

```{r}
write_sf(pop_region_3338, "data/ak_regions_pop.shp", delete_layer = TRUE)# delete layer it removes previous
```

## Make maps!

### using scale package to change labels to comma

```{r}
ggplot() +
  geom_sf(data = pop_region_3338, aes(fill = total_pop)) + 
  theme_bw() +
  labs(fill = "Total Population") +
  scale_fill_continuous(low = "khaki", high = "firebrick", labels = comma) #labels comes from scale packages. 
```

### adding points
```{r}
ggplot() +
  geom_sf(data = pop_region_3338, aes(fill = total_pop)) + 
  geom_sf(data = pop_3338, aes(), size = 0.5) +
  theme_bw() +
  labs(fill = "Total Population") +
  scale_fill_continuous(low = "khaki", high = "firebrick", labels = comma) #labels comes from scale packages. 
```

### adding rivers 

```{r}
rivers_3338 <- read_sf("data/ak_rivers_simp.shp")

st_crs(rivers_3338)

ggplot() +
  geom_sf(data = pop_region_3338, aes(fill = total_pop)) + 
  geom_sf(data = rivers_3338, aes(size = StrOrder), color = "black") +
  geom_sf(data = pop_3338, aes(), size = 0.5) +
  scale_size(range = c(0.01, 0.2), guide = F) + #to avoid problems with different scales
  theme_bw() +
  labs(fill = "Total Population") +
  scale_fill_continuous(low = "khaki", high = "firebrick", labels = comma) #labels comes from scale packages. 
```

### Getting basemaps using ggmap it works with 3857

```{r}
pop_3857 <- pop_3338 %>% 
  st_transform(crs = 3857)
```

We are getting a stamenmap from stamenmap
```{r}
# Define a function to fix the bbox to be in EPSG:3857
# See https://github.com/dkahle/ggmap/issues/160#issuecomment-397055208
ggmap_bbox_to_3857 <- function(map) {
  if (!inherits(map, "ggmap")) stop("map must be a ggmap object")
  # Extract the bounding box (in lat/lon) from the ggmap to a numeric vector, 
  # and set the names to what sf::st_bbox expects:
  map_bbox <- setNames(unlist(attr(map, "bb")), 
                       c("ymin", "xmin", "ymax", "xmax"))
  
  # Coonvert the bbox to an sf polygon, transform it to 3857, 
  # and convert back to a bbox (convoluted, but it works)
  bbox_3857 <- st_bbox(st_transform(st_as_sfc(st_bbox(map_bbox, crs = 4326)), 3857))
  
  # Overwrite the bbox of the ggmap object with the transformed coordinates 
  attr(map, "bb")$ll.lat <- bbox_3857["ymin"]
  attr(map, "bb")$ll.lon <- bbox_3857["xmin"]
  attr(map, "bb")$ur.lat <- bbox_3857["ymax"]
  attr(map, "bb")$ur.lon <- bbox_3857["xmax"]
  map
}
```

```{r}
bbox <- c(-170, 52, -130, 64)

ak_map <- get_stamenmap(bbox, zoom = 4)

ak_map_3857 <- ggmap_bbox_to_3857(ak_map)

class(ak_map_3857)
```

Mapping
```{r}
ggmap(ak_map_3857) + 
  geom_sf(data = pop_3857, aes(color = population), inherit.aes = F) +
  scale_color_continuous(low = "khaki", high =  "firebrick", labels = comma)
```

## Interactive mapping using leaflet

Here we define a leaflet projection for Alaska Albers, and save it as a variable to use later.

```{r}
epsg3338 <- leaflet::leafletCRS(
  crsClass = "L.Proj.CRS",
  code = "EPSG:3338",
  proj4def =  "+proj=aea +lat_1=55 +lat_2=65 +lat_0=50 +lon_0=-154 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs",
  resolutions = 2^(16:7))
```

let’s use st_transform yet again to get back to WGS84

```{r}
pop_region_4326 <- pop_region_3338 %>% 
  st_transform(crs = 4326)
```

generate the map

```{r}
pal <- colorNumeric(palette = "Reds", domain = pop_region_4326$total_pop)

m <- leaflet(options = leafletOptions(crs = epsg3338)) %>%
        addPolygons(data = pop_region_4326, 
                    fillColor = ~pal(total_pop),
                    weight = 1,
                    color = "black",
                    fillOpacity = 1,
                    label = ~region) %>% 
        addLegend(position = "bottomleft",
                  pal = pal,
                  values = range(pop_region_4326$total_pop),
                  title = "Total Population")
m
```

with ti
```{r}
m <- leaflet() %>%
  addTiles() %>% 
        addPolygons(data = pop_region_4326, 
                    fillColor = ~pal(total_pop),
                    weight = 1,
                    color = "black",
                    fillOpacity = 1,
                    label = ~region) %>% 
        addLegend(position = "bottomleft",
                  pal = pal,
                  values = range(pop_region_4326$total_pop),
                  title = "Total Population")
m
```

